---
title: "Intensity Functional Principal Component Analysis"
output: github_document
---

## Overview

This package re-implements the first step of the {[MASTA](https://celehs.github.io/MASTA/)} package to extract features from longitudinal encounter records. Compared to {MASTA}, the input data of {IFPCA} is more compact and memory efficent. Click [HERE](https://github.com/celehs/IFPCA/tree/master/data-raw) to view input data structure.

## Installation

Install development version from GitHub.

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("celehs/IFPCA")
```

Load the package into R.

```{r}
library(IFPCA)
```

## Data Example

### Data Preparation

```{r}
library(data.table)
```

```{r}
i <- 3 # 1, 2, 3 
# time for each code (training + validation)
time_code <- fread(paste0("data-raw/time_code", i, ".csv"))
time <- time_code$month
names(time) <- time_code$id
# follow up time for training and validation sets
follow_up_train <- fread("data-raw/follow_up_train.csv")
follow_up_valid <- fread("data-raw/follow_up_valid.csv")
fu_train <- follow_up_train$fu_time
fu_valid <- follow_up_valid$fu_time
names(fu_train) <- follow_up_train$id
names(fu_valid) <- follow_up_valid$id
str(fu_train)
str(fu_valid)
```

## Itensity FPCA

```{r}
system.time(ans <- ifpca(time, fu_train, fu_valid))
```

```{r}
data.table(ans$TrainFt) # Extracted Features (Training) 
data.table(ans$ValidFt) # Extracted Features (Validation)
```

## References

- Wu, S., Müller, H., Zhang, Z. (2013). __Functional Data Analysis for Point Processes with Rare Events__. _Statistica Sinica_, 23:1-23. <https://doi.org/10.5705/ss.2010.162>

- Liang, L., Uno, H., Ma, Y., Cai, T. __Robust Approach to Event Time Annotation Using Longitudinal Medical Encounters__. _Working Paper_.
